---
title: "Linear Regression"
author: "Mariano, Roshni and Pedram"
date: "9/22/2018"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette summarises functions of LR package and gives examples on how to use them.



#DATA
To explore the basic functions of LR, we'll use ``` iris ```.This is perhaps the best known database to be found in the pattern recognition literature. The data set contains 3 classes of 50 instances each, where each class refers to a type of iris plant. One class is linearly separable from the other 2; the latter are NOT linearly separable from each other. you can see first 10 rows below:
```{r,echo=FALSE}
head(iris,10)
```


#Included functions 
LR has a vary of functions to provide analasis on data.


    *print()   prints out the coeffcients and coeffcient names.
    *plot()    plots  two plots using ggplot2
    *resid()   returns the vector of residuals e.
    *pred()    returns the predicted values ^y.
    *coef()    returns the coeffcients as a named vector.
    *summary() returns a similar printout as printed for lm objects























## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
