---
title: "Linear Regression"
author: "Mariano, Roshni and Pedram"
date: "9/22/2018"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r,echo=FALSE}

linreg <- setRefClass(Class = "linreg",
                      
                    
                      fields = list(formula="formula", data="data.frame", regco="matrix", yf="matrix", e="matrix", dfreedom="numeric", 
                                    Sigma_square="numeric", Var_Beta="matrix", t_Beta="matrix", pvalue="matrix",
                                    parse="character", stand_res="matrix",variance="numeric"),
                      
                      
                      
                      methods = list(
                        
                        initialize =function (formula,data)
                        {
                          c<-colnames(data)
                          d<-all.vars(formula)
                          stopifnot(d %in% c)
                          stopifnot (is.data.frame(data))
                          formula <<- formula
                          data <<- data
                          X <- model.matrix(formula,data)
                          dep_y <- all.vars(formula)[1]
                          y <- (data[,dep_y])
                          parse <<- deparse(substitute(data))
                          #Regressions coefficients
                          regco <<- solve((t(X)%*%X))%*%t(X)%*%y
                          #X <- QR
                          #Beta <- solve(R)%*%t(Q)%*%y
                          #Fitted values
                          yf <<- X%*%regco
                          #Residuals
                          e <<- y-yf
                          #Degrees of freedom
                          dfreedom <<- nrow(X)-ncol(X)
                          #Residual variance
                          Sigma_square <<- as.numeric((t(e)%*%e) / dfreedom)
                          #Variance of regression coefficients
                          Var_Beta <<- Sigma_square * solve((t(X)%*%X))
                          #t-values for each coefficient
                          t_Beta <<- regco / sqrt(diag(Var_Beta))
                          #p values for reg coefficients
                          pvalue <<- pt(abs(t_Beta),dfreedom)
                          #variance value
                          variance <<- round(sqrt(Sigma_square),2)
                          #standardised residual for plot2
                          stand_res <<- sqrt(abs((e-mean(e)) / sqrt(Sigma_square)))
                        
                        },
                        #Methods
                        #print(),plot(), resid(),pred(),coef(),summary()
                        print = function(){
                          regco
                          
                        },
                        
                        
                        plot = function(){
                          library(ggplot2)
                          #plotting yf and e
                          data_frame1 <- data.frame(Fitted_values=yf,Residuals=e)
                          p1 <- ggplot(data_frame1,aes(Fitted_values,Residuals))+
                            geom_point()+geom_abline()+
                            ggtitle("Residuals vs Fitted")+
                            xlab("Fitted values lm(Petal.Length ~ Species)")+
                            ylab("Residuals")
                          
                          data_frame2 <- data.frame(Fitted_values=yf,Residuals=stand_res)
                          p2 <- ggplot(data_frame2,aes(Fitted_values,Residuals))+
                            geom_point()+geom_abline()+
                            ggtitle("Scale-Location")+
                            xlab("Fitted values lm(Petal.Length ~ Species)")+
                            ylab(expression(bold(sqrt(bold("Standardized Residuals")))) )
                          return(list(p1, p2))
                        },
                        
                        resid = function(){
                          cat("Returning vector of residuals e:", "\n")
                          return(as.vector(round(e,2)))
                        },
                        
                        pred = function(){
                          cat("Returning predicted values yf:", "\n")
                          return(as.vector(round(yf,2)))
                        },
                        
                        coef = function(){
                          cat("Returning coefficients as a vector:", "\n")
                          return(as.vector(round(Beta,2)))
                        },
                        
                        summary = function(){
                          
                          base::summary()
                          
                          
                        }
                        
                        
                        
                      ))

```

```{r,echo=FALSE}
data <- iris
linreg <- function (formula,data)
                        {
                          c<-colnames(data)
                          d<-all.vars(formula)
                          stopifnot(d %in% c)
                          stopifnot (is.data.frame(data))
                          formula <<- formula
                          data <<- data
                          X <- model.matrix(formula,data)
                          dep_y <- all.vars(formula)[1]
                          y <- (data[,dep_y])
                          parse <<- deparse(substitute(data))
                          #Regressions coefficients
                          regco <<- solve((t(X)%*%X))%*%t(X)%*%y
                          #X <- QR
                          #Beta <- solve(R)%*%t(Q)%*%y
                          #Fitted values
                          yf <<- X%*%regco
                          #Residuals
                          e <<- y-yf
                          #Degrees of freedom
                          dfreedom <<- nrow(X)-ncol(X)
                          #Residual variance
                          Sigma_square <<- as.numeric((t(e)%*%e) / dfreedom)
                          #Variance of regression coefficients
                          Var_Beta <<- Sigma_square * solve((t(X)%*%X))
                          #t-values for each coefficient
                          t_Beta <<- regco / sqrt(diag(Var_Beta))
                          #p values for reg coefficients
                          pvalue <<- pt(abs(t_Beta),dfreedom)
                          #variance value
                          variance <<- round(sqrt(Sigma_square),2)
                          #standardised residual for plot2
                          stand_res <<- sqrt(abs((e-mean(e)) / sqrt(Sigma_square)))
                        
                        }
```

This vignette summarizes functions of LR package and gives examples on how to use them.



#DATA
To explore the basic functions of LR, we'll use ``` iris ```.This is perhaps the best known database to be found in the pattern recognition literature. The data set contains 3 classes of 50 instances each, where each class refers to a type of iris plant. One class is linearly separable from the other 2; the latter are NOT linearly separable from each other. you can see first 10 rows below:
```{r,echo=FALSE}
head(iris,10)
```


#Included functions 
LR has a vary of functions to provide analysis  on data.


    *print()   prints out the coefficients and coefficient names.
    *plot()    plots  two plots using ggplot2
    *resid()   returns the vector of residuals e.
    *pred()    returns the predicted values ^y.
    *coef()    returns the coefficients as a named vector.
    *summary() returns a similar printout as printed for lm objects
    

##Print()


prints out the coefficients and coefficient names, similar as done by the lm class.

```{r,echo=FALSE}
data(iris)
mod_object <- lm(Petal.Length~Species, data = iris)
print(mod_object)
```

##Plot()

This function plots two plots using ggplot2 shown below:

```{r,echo=FALSE}
 library(ggplot2)
data <- iris
formula = Petal.Length ~ Species
                          c<-colnames(data)
                          d<-all.vars(formula)
                          stopifnot(d %in% c)
                          stopifnot (is.data.frame(data))

                          X <- model.matrix(formula,data)
                          dep_y <- all.vars(formula)[1]
                          y <- (data[,dep_y])
                          parse<- deparse(substitute(data))
                          #Regressions coefficients
                          regco <<- solve((t(X)%*%X))%*%t(X)%*%y
                          #X <- QR
                          #Beta <- solve(R)%*%t(Q)%*%y
                          #Fitted values
                          yf <<- X%*%regco
                          #Residuals
                          e <<- y-yf
                          #Degrees of freedom
                          dfreedom <<- nrow(X)-ncol(X)
                          #Residual variance
                          Sigma_square <<- as.numeric((t(e)%*%e) / dfreedom)
                          #Variance of regression coefficients
                          Var_Beta <<- Sigma_square * solve((t(X)%*%X))
                          #t-values for each coefficient
                          t_Beta <<- regco / sqrt(diag(Var_Beta))
                          #p values for reg coefficients
                          pvalue <<- pt(abs(t_Beta),dfreedom)
                          #variance value
                          variance <<- round(sqrt(Sigma_square),2)
                          #standardised residual for plot2
                          stand_res <<- sqrt(abs((e-mean(e)) / sqrt(Sigma_square)))
                        
                          #plotting yf and e
                          data_frame1 <- data.frame(Fitted_values=yf,Residuals=e)
                          p1 <- ggplot(data_frame1,aes(Fitted_values,Residuals))+
                            geom_point()+geom_abline()+
                            ggtitle("Residuals vs Fitted")+
                            xlab("Fitted values lm(Petal.Length ~ Species)")+
                            ylab("Residuals")
                          
                          data_frame2 <- data.frame(Fitted_values=yf,Residuals=stand_res)
                          p2 <- ggplot(data_frame2,aes(Fitted_values,Residuals))+
                            geom_point()+geom_abline()+
                            ggtitle("Scale-Location")+
                            xlab("Fitted values lm(Petal.Length ~ Species)")+
                            ylab(expression(bold(sqrt(bold("Standardized Residuals")))) )
                          return(list(p1, p2))
```


#resid()
this function returns the vector of residuals e.
```{r,echo=FALSE}

resid(cat("Returning vector of residuals e:", "\n"))
                          return(as.vector(round(e,2)))
```


#pred()
returns the  predicted values of y_hat.
```{r,echo=FALSE}
             pred = function(){
                          cat("Returning predicted values yf:", "\n")
                          return(as.vector(round(yf,2)))
                        }
                          return(as.vector(round(yf,2)))

```



#coef()

This function returns the cofficients as a named vector
```{r,echo=FALSE}
                        coef = function(){
                          cat("Returning coefficients as a vector:", "\n")
                          return(as.vector(round(regco,2)))
                        }
return(as.vector(round(regco,2)))
```


#summary()


this function returns a similar printout as printed for lm objects, but you only need to present the coefficients with their standard error, t-value and p-value as well as the estimate of sigma and the degrees
of freedom in the model.

















